### 2023-09-23 appendices/paper_brief.tex
Work on variational autoencoders

\chapter[a brief overview of unsupervised neural speech representation learning]{A Brief Overview of Unsupervised Neural Speech Representation Learning}
\label{app:paper-brief}
% \ifthenelse{\equal{\skipappendices}{true}}{}{

% \newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}

% \newcommand{\norm}[1]{\left\lVert#1\right\rVert}

@@

\section{Introduction}
Representation learning has shaped modern computer vision \cite{simonyan2014very} and natural language processing \cite{devlin2019bert}, and more recently speech processing has been subject to the same development \cite{baevski2020wav2vec}. Representation learning has been defined as \emph{``learning representations of the data that make it easier to extract useful information when building classifiers or other predictors"} \cite{bengio_representation_2013}.  
Unsupervised representation learning is concerned with learning useful representations without the use of human annotations. Usually, a model is first pre-trained on a task where plenty of data is available. The model is then fine-tuned, or used to extract input representations for a smaller model, targeting a task with limited training data. In computer vision, both supervised \cite{simonyan2014very, szegedy2015going, he2016deep} and unsupervised \cite{pathak2016context, doersch2015unsupervised} representation learning have gained attention with supervised representation learning driven by the availability of large annotated datasets \cite{deng2009imagenet}. For text and speech, pre-training is usually unsupervised as labeled data is difficult to obtain. Although work on text has paved the way, and the two fields share many characteristics, learning representations from speech is a problem faced with a unique set of challenges.

In this paper, we survey work on unsupervised representation learning for speech processing from within the last decade.
From a methodological perspective, we identify two primary model categories, namely models based on self-supervised learning and probabilistic latent variable models. 
We provide a methodological review of the design choices related to each of the model categories and develop a model taxonomy that highlights the different directions of work. Finally, we compare and discuss models from the two categories and their respective evaluation procedures.
%We then provide a comparison of the methods based on their respective evaluation procedures.
%Finally, based on the model taxonomy, we discuss the differences between models from the two categories and identify future challenges and directions for research in speech representation learning.

@@

\section{Unsupervised representation learning}

@@

%Similarly, $\textbf{a}_{<t}$ denotes a vector-sequence containing all elements prior to step $t$ and $\textbf{a}_{t \in \mathbb{A}}$ is a vector sequence containing elements with index contained in the set $\mathbb{A}$.
%(or a scalar sequence $a_{i:j}$)

%Input to a model is either a spectrogram $\mathbf{x}_{1:T}$ or the raw speech signal $x_{1:L}$ where $L > T$. Commonly $L = T(\frac{S}{100})$ where $S$ is the sample rate of the raw speech signal.

% \IEEEpubidadjcol % related to the IEEEpubid command in the top

\paragraph{Notation} \label{ssec:notation} We use the subscript $i{:}j$ with $i{\leq}j$ to denote a vector sequence $\textbf{a}_{i:j}$ containing elements $\textbf{a}_{i}$ through $\textbf{a}_{j}$. We denote model input as $\mathbf{x}_{1:T}$ which, in practice, might be either a spectrogram or the raw speech signal, but we do not distinguish between the two in notation as it is not essential to understand the models. Also, models commonly downsample the temporal dimension, but again, this is not crucial to understand the models, so we maintain a notation based on a single temporal dimension $t\in\{1,\dots, T\}$.

When discussing self-supervised models, we use $\mathbf{c}_{1:T}$ to denote a contextualized representation. For stochastic latent variable models, we use $\mathbf{z}_{1:T}$ as is customary to the field. 
% The representation at a single time-step, $\mathbf{c}_t$ or $\mathbf{z}_t$, is typically a function of a larger part of the input space than just $\mathbf{x}_t$, in some cases the entire input sequence $\mathbf{x}_{1:T}$.
While some models are frozen and produce representations used as input for downstream tasks (\textbf{\textsc{frz}}, table \ref{tab:model-taxonomy}), others are designed to be fine-tuned (\textbf{\textsc{ftn}}, table \ref{tab:model-taxonomy}). In either case, we use $f(\cdot)$ to denote the model that is used for the downstream task. We use $g(\cdot)$ to denote any auxiliary model components (e.g., for a reconstruction task we might have $g: \mathbf{c}_t \mapsto \hat{\mathbf{x}}_t$). When a model can be naturally subdivided into multiple components, we simply use $f_*(\cdot)$ where $*$ may be any convenient descriptor. Finally, we often use a subscript when defining a loss, $\mathcal{L}_i$, to imply that the total loss is computed as a sum over $i$.

@@

Self-supervised learning is a subset of unsupervised learning \cite{tsai2020self}. Where other unsupervised methods can be seen as a means to an end in itself (e.g., clustering or data generation), self-supervised learning takes the form of a pretext task that only adds value when associated with a downstream task. This makes self-supervised learning tie naturally with semi-supervised learning, but it may also be part of a fully unsupervised setup \cite{baevski2021unsupervised}. Self-supervised learning is often characterized by automatically deriving the target from the input or other unlabeled examples \cite{ouali2020overview}.

\paragraph{Predictive models} Similar to classic autoregressive language models \cite{mikolov2010recurrent}, contextualized speech representations can be learned by predicting future values of a simple representation \cite{oord2018representation, chung2019unsupervised, schneider2019wav2vec, chung2020generative, jiang2021further} (\textbf{\textsc{prd}}, table \ref{tab:model-taxonomy}). Modeling spectrograms directly, autoregressive predictive coding (APC, \citealp{chung2019unsupervised}) is perhaps the simplest example in this category. The forward pass and loss are computed as
\begin{align}
    \mathbf{c}_{t} &= f(\mathbf{x}_{1:t})  \label{eq:apc_f} \\ 
    \hat{\mathbf{x}}_{t+k} &= g(\mathbf{c}_{t}) \\
    %\mathcal{L}_t(\mathbf{\hat{x}}_{t+k}, \mathbf{x}_{t+k}) &= |\mathbf{\hat{x}}_{t+k} - \mathbf{x}_{t+k}|
    \mathcal{L}_t &= \lVert \hat{\mathbf{x}}_{t+k} - \mathbf{x}_{t+k} \rVert_1\enspace.
\end{align}
%
\noindent Here, $f(\cdot)$ and $g(\cdot)$ are parameterized by neural networks such that each $\mathbf{c}_t$ is only conditioned on previous inputs $\mathbf{x}_{1:t}$ and $\hat{\mathbf{x}}_{t+k}$ is computed step-wise. \citet{chung2019unsupervised} use a stack of unidirectional LSTMs for $f(\cdot)$ and a linear regression layer for $g(\cdot)$. Tasks that seek to predict or reconstruct the input are very common. In the literature, these are often jointly referred to as reconstruction tasks (\textbf{\textsc{rec}}, table \ref{tab:model-taxonomy}) \cite{liu2021tera, wang2021unispeech}, although this is somewhat misleading in the case of prediction.

Contrary to generative models, such as WaveNet \cite{oord_wavenet:_2016}, the APC model is not restricted to next-step prediction. Instead, it predicts $k > 0$ steps ahead in order to ensure that the model does not learn a trivial solution by exploiting the smoothness of the signal. Depending on the downstream task, we are often interested in learning so-called slow features that will typically span multiple input frames \cite{wiskott2002slow}. Even the smallest linguistic units of speech, phonemes, tend to span $0.1$ seconds on average \cite{garofolo_timit_1993}, whereas spectrogram frames $\mathbf{x}_t$ are typically computed at $0.01$ second intervals. However, sometimes local smoothness is explicitly used to define the task \cite{badino2014auto, jati2017speaker2vec, jati2019neural}.

%In early work, \citet{badino2014auto} propose a step-wise bottleneck model that learns to predict $\mathbf{x}_{t+1}$ from $\mathbf{x}_{t}$. Since phonetic information changes slowly, $\mathbf{x}_{t+1}$ is argued to be a noisy version of $\mathbf{x}_{t}$ encoding almost the same phonetic content. Thus, the model is akin to a denoising autoencoder.
%A similar idea has been proposed by Jati and Georgiou \cite{jati2019neural} to learn speaker embeddings.

% Next-step prediction has also been frequently used for generative models (\textbf{\textsc{gen}}, table \ref{tab:model-taxonomy}). The popular speech synthesis model WaveNet is perhaps the most obvious example \cite{oord_wavenet:_2016}. While this work does not consider representation learning, the model might learn useful features similar to those learned with APC.

\paragraph{Contrastive models} Speech contains localized noise (e.g., phase shifts) that does not inform slow feature learning. Thus, directly modeling speech might not be the best way to learn contextualized representations. Contrastive predictive coding (CPC,  \citealp{oord2018representation}) targets a local variable $\mathbf{v}_{1:T}$, learned from the model input $\mathbf{x}_{1:T}$, instead of the input itself. The forward pass is
%
\begin{align}
    \mathbf{v}_{t} &= f_{v}(\mathbf{x}_{t-r:t+r}) \label{eq: cpc local representation} \\
    \mathbf{c}_{t} &= f_{c}(\mathbf{v}_{1:t}) \\
    \hat{\mathbf{v}}_{t,k} &= g_k(\mathbf{c}_{t}) \enspace,
    %g_k(\mathbf{c}_{t}, \mathbf{v}_{t+k}) &= \mathbf{v}_{t+k}^{\intercal} \mathbf{W}_k \mathbf{c}_{t}
    %\mathcal{L}_{t,k} &= - \log \left(\frac{g_k(\mathbf{c}_{t}, \mathbf{v}_{t+k})}{\sum_{n \sim \mathcal{D}} g_k(\mathbf{c}_{t}, \mathbf{v}_{n})} \right)
    % \mathcal{L}_{t,k} &= -\log g_k + \log \sum_{n \sim \mathcal{D}} g_k
\end{align}
%
\noindent where $f_v(\cdot)$ is a convolutional neural network, such that each $\mathbf{v}_{t}$ only encodes information from a limited receptive field $2r+1$. Again, $f_c(\cdot)$ should be limited to condition each $\mathbf{c}_t$ on previous time-steps $\mathbf{v}_{1:t}$ and $g_k(\cdot)$ is a step-wise transformation. The loss is based on noise constrastive estimation \cite{gutmann2010noise} and is given by
\begin{align}
    \mathcal{L}_{t,k} &= - \log \left(\frac{\exp(\hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{t+k})}{\sum_{n \sim \mathcal{D}} \exp(\hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{n})} \right)\enspace .
    % \mathcal{L}_{t,k} &= - \hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{t+k} + \log \sum_{n \sim \mathcal{D}} \exp(\hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{n})\enspace .
    % \mathcal{L}_{t,k} &= -\log g_k + \log \sum_{n \sim \mathcal{D}} g_k
    \label{eq: cpc loss}
\end{align}
%
\noindent Here, $\mathcal{D}$ is a set of indices including the target index $t+k$ and negative samples drawn from a proposal distribution, which is typically taken to be a uniform distribution over the set $\{1,\dots,T\}$. Note that the loss is also indexed by $k$ to show that CPC targets multiple offsets. The APC model is easily extended in a similar way \cite{chung2020improved}.

@@

A model closely related to the original CPC model is wav2vec \cite{schneider2019wav2vec}. It uses a different parameterization of the functions $f_v(\cdot)$ and $f_c(\cdot)$, and modifies the loss to consider a binary prediction task, such that we have
%
\begin{align}
    %g_k(\mathbf{c}_{t}, \mathbf{v}_{t+k}) &= \log(\sigma(\mathbf{v}_{t+k}^{\intercal} (\mathbf{W}_k \mathbf{c}_{t} + \mathbf{b}_k)))\\
    \mathcal{L}_{t,k} &= - \log(\sigma(\hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{t+k})) - \sum_{n \sim \mathcal{D}} \log(\sigma(-\hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{n})) \,.
    \label{eq: wav2vec loss}
\end{align}

@@

\paragraph{Masking-based models} One downside of predictive tasks is that models are primarily unidirectional. Some work has extended APC and CPC inspired models with separate encoders operating in opposite directions \cite{ling2020deep, kawakami2020learning, borgholt2021scaling}, but these models are still restricted to process left and right context separately. Inspired by the masked language model task used for text-based representation learning \cite{devlin2019bert}, several papers have used masking to overcome this challenge (\textbf{\textsc{msk}}, table \ref{tab:model-taxonomy}). Masking refers to replacing parts of the input with zeros or a learned masking vector. For zero-masking \cite{jiang2019improving, liu2020mockingjay, wang2020unsupervised, chi2021audio, ling2020decoar}, we have
\begin{align}
    \mathbf{c}_{1:T} &= f(\mathbf{x}_{1:T} \circ \mathbf{m}_{1:T})\\
    \mathbf{\hat{x}}_{t} &= g(\mathbf{c}_{t}) \\
    \mathcal{L}_t &= \lVert \mathbf{\hat{x}}_{t} - \mathbf{x}_{t} \rVert_1\enspace,
\end{align}
where the $\circ$ operator denotes the Hadamard product, $f(\cdot)$ is typically a transformer encoder or a bidirectional recurrent neural network, $g(\cdot)$ is a step-wise transformation, and $\mathbf{m}_{1:T}$ is a mask such that $m_{t,i} \in \{0,1\}$. Alternatively, $\mathbf{m}_{1:T}$ is used to select which $\mathbf{x}_t$ are replaced by a learned masking vector. The entries of $\mathbf{m}_{1:T}$ are determined by some stochastic policy. One frequent inspiration is SpecAugment \cite{park2019specaugment}, which was originally proposed for supervised speech recognition and applies frequency and time masking to spectrogram representations.
While temporal masking is most common, frequency masking has also been adopted for representation learning  \cite{wang2020unsupervised}. A simple, yet popular, masking strategy is to draw a proportion of input indices $t_i\sim\{1,\dots,T-M\}$ without replacement, and then mask $\{t_i, \dots, t_i+M\}$ \cite{baevski2020wav2vec, hsu2021hubert, ling2020decoar}.

Combining masking with a contrastive loss, wav2vec 2.0 was the first work to show that a competitive speech recognition model can be learned by fine-tuning a pre-trained model with as little as 10 minutes of labeled data. For this model
%
\begin{align}
    \mathbf{v}_{t} &= f_v(\mathbf{x}_{t-r:t+r}) \\
    \mathbf{c}_{1:T} &= f_c(\mathbf{v}_{1:T} \circ \mathbf{m}_{1:T}) \\
    \mathbf{q}_t &= g_q(\mathbf{v}_t) \enspace . \label{w2v2 qtz}
\end{align}

\noindent Here, $f_v(\cdot)$ is a convolutional neural network, $f_c(\cdot)$ is a transformer encoder \cite{vaswani2017attention} and $g_q(\cdot)$ is a quantization module used to learn targets from the localized variable $\mathbf{v}_{1:T}$. Computing quantized targets this way requires an extra loss term, which we will present when we discuss quantization in general below. The contrastive loss for wav2vec 2.0 is similar to that of the CPC model,
%
\begin{align}
    \mathcal{L}_t &= - \log \left(\frac{\exp(S_{\text{c}}(\mathbf{c}_{t}, \mathbf{q}_{t}))}{\sum_{n \sim \mathcal{D}} \exp(S_{\text{c}}(\mathbf{c}_{t}, \mathbf{q}_{n}))} \right) \enspace , \label{w2v2 loss}
\end{align}
%
\noindent where $S_{\text{c}}(\cdot)$ is the cosine similarity and the negative samples in $\mathcal{D}$ are sampled from other masked time-steps.

@@

\noindent Non-autoregressive predictive coding (NPC, \citealp{liu2020non}) tries to resolve this by using a convolutional neural network where the kernel is masked instead of the input.
%That is, the model is prevented from using $\mathbf{x}_{t}$ for its corresponding reconstruction $\mathbf{\hat{x}}_{t}$.
This allows for complete data utilization, but limits the amount of context encoded in the learned representation.
Figure \ref{fig:ssl_grid} summarizes the models discussed so far.

\begin{figure}[!t]
    \centering
    \setlength\tabcolsep{1.5pt}
    \begin{tabular}{>{\centering\arraybackslash} m{4mm}  >{\centering\arraybackslash} m{0.45\textwidth}|>{\centering\arraybackslash} m{0.45\textwidth}}
          & {\small \textsc{predictive}} & {\small \textsc{masking-based}} \\
        %\midrule
        \rotatebox{90}{{\small \textsc{reconstruct}}} & \includegraphics[width=0.45\textwidth]{paper_brief/REC_PRD.pdf} & \includegraphics[width=0.45\textwidth]{paper_brief/REC_MSK.pdf}  \\
        \midrule
        \rotatebox{90}{{\small \textsc{contrastive}}} & \includegraphics[width=0.45\textwidth]{paper_brief/CON_PRD.pdf} & \includegraphics[width=0.45\textwidth]{paper_brief/CON_MSK.pdf}
    \end{tabular}
    \caption{
    Schematic of self-supervised methods. Each subfigure illustrates the loss computation for a single time-step.
    The temporal subscript has been left out for simplicity. 
    %Examples of models in each category include APC (\textbf{\textsc{prd}} + \textbf{\textsc{rec}}), Mockingjay (\textbf{\textsc{msk}} + \textbf{\textsc{rec}}), CPC (\textbf{\textsc{prd}} + \textbf{\textsc{con}}) and wav2vec 2.0 (\textbf{\textsc{msk}} + \textbf{\textsc{con}}).
    }
    \label{fig:ssl_grid}
\end{figure}

@@

            \midrule
            \multicolumn{12}{c}{\textsc{Self-supervised models}} \\
            \midrule

            %                                                           MSK      PRD      CON      REC      QTZ      GEN      LOC      GLB     VAR      FRZ      FTN
            \textbf{Audio Word2vec} \footnotesize\cite{chung2016audio}      & 2016 Mar. & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark \\
            \textbf{Speech2Vec} \footnotesize\cite{chung2018speech2vec}     & 2018 Mar. & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark \\
            \textbf{Unspeech} \footnotesize\cite{milde2018unspeech}         & 2018 Apr. & \xmark & \cmark & \cmark & \xmark & \xmark & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark \\

@@

The inference model may also be used to infer a global latent variable (\textbf{\textsc{glb}}, table \ref{tab:lvm-model-primitives}) that might encode global information about $\mathbf{x}$. 
While it must be included in the prior model it might not be in the observation model, if the model also has a local latent variable. 
Finally, latent variables are often made to depend autoregressively on past inferred values, e.g. $q(\mathbf{z}_t|\mathbf{z}_{1:t-1},\mathbf{x}_{1:t})$ (\textbf{\textsc{arz}}, table \ref{tab:lvm-model-primitives}) \cite{chung_recurrent_2015, fraccaro_sequential_2016}.
% Finally, latent variables are often made to depend autoregressively on past values, $q(\mathbf{z}_t|\mathbf{z}_{1:t-1},\mathbf{x}_{1:t})$ for filtering and $q(\mathbf{z}_t|\mathbf{z}_{1:t-1},\mathbf{x}_{1:T})$ for smoothing (\textbf{\textsc{arz}}, table \ref{tab:lvm-model-primitives}) \cite{chung_recurrent_2015, fraccaro_sequential_2016}.
%Such autoregressive dependencies help the inference model better match the prior dynamics but notably are not employed in models that learn the prior dynamics post-hoc \cite{oord_neural_2018}.

>>>>

\chapter[a brief overview of unsupervised neural speech representation learning]{A Brief Overview of Unsupervised Neural Speech Representation Learning}
\label{app:paper-brief}
\ifthenelse{\equal{\skipappendices}{true}}{}{

@@

\section{Introduction}
Representation learning has shaped modern computer vision \cite{simonyan_very_2014} and natural language processing \cite{devlin_bert_2018}, and more recently speech processing has been subject to the same development \cite{baevski_wav2vec_2020}. Representation learning has been defined as \emph{``learning representations of the data that make it easier to extract useful information when building classifiers or other predictors"} \cite{bengio_representation_2013}.  
Unsupervised representation learning is concerned with learning useful representations without the use of human annotations. Usually, a model is first pre-trained on a task where plenty of data is available. The model is then fine-tuned, or used to extract input representations for a smaller model, targeting a task with limited training data. In computer vision, both supervised \cite{simonyan_very_2014, szegedy_going_2015, he_deep_2015} and unsupervised \cite{pathak_context_2016, doersch_unsupervised_2015} representation learning have gained attention with supervised representation learning driven by the availability of large annotated datasets \cite{deng_imagenet_2009}. For text and speech, pre-training is usually unsupervised as labeled data is difficult to obtain. Although work on text has paved the way, and the two fields share many characteristics, learning representations from speech is a problem faced with a unique set of challenges.

In this paper, we survey work on unsupervised representation learning for speech processing from within the last decade.
From a methodological perspective, we identify two primary model categories, namely models based on self-supervised learning and probabilistic latent variable models. 
We provide a methodological review of the design choices related to each of the model categories and develop a model taxonomy that highlights the different directions of work. Finally, we compare and discuss models from the two categories and their respective evaluation procedures.

@@

\section{Unsupervised representation learning}\label{ssec:notation}

@@

\paragraph{Notation}
We use the subscript $i{:}j$ with $i{\leq}j$ to denote a vector sequence $\textbf{a}_{i:j}$ containing elements $\textbf{a}_{i}$ through $\textbf{a}_{j}$. We denote model input as $\mathbf{x}_{1:T}$ which, in practice, might be either a spectrogram or the raw speech signal, but we do not distinguish between the two in notation as it is not essential to understand the models. Also, models commonly downsample the temporal dimension, but again, this is not crucial to understand the models, so we maintain a notation based on a single temporal dimension $t\in\{1,\dots, T\}$.

When discussing self-supervised models, we use $\mathbf{c}_{1:T}$ to denote a contextualized representation. For stochastic latent variable models, we use $\mathbf{z}_{1:T}$ as is customary to the field. 
While some models are frozen and produce representations used as input for downstream tasks (\textbf{\textsc{frz}}, table \ref{tab:model-taxonomy}), others are designed to be fine-tuned (\textbf{\textsc{ftn}}, table \ref{tab:model-taxonomy}). In either case, we use $f(\cdot)$ to denote the model that is used for the downstream task. We use $g(\cdot)$ to denote any auxiliary model components (e.g., for a reconstruction task we might have $g: \mathbf{c}_t \mapsto \hat{\mathbf{x}}_t$). When a model can be naturally subdivided into multiple components, we simply use $f_*(\cdot)$ where $*$ may be any convenient descriptor. Finally, we often use a subscript when defining a loss, $\mathcal{L}_i$, to imply that the total loss is computed as a sum over $i$.

@@

Self-supervised learning is a subset of unsupervised learning \cite{tsai_selfsupervised_2021}. Where other unsupervised methods can be seen as a means to an end in itself (e.g., clustering or data generation), self-supervised learning takes the form of a pretext task that only adds value when associated with a downstream task. This makes self-supervised learning tie naturally with semi-supervised learning, but it may also be part of a fully unsupervised setup \cite{baevski_unsupervised_2021}. Self-supervised learning is often characterized by automatically deriving the target from the input or other unlabeled examples \cite{ouali_overview_2020}.

\paragraph{Predictive models}
Similar to classic autoregressive language models \cite{mikolov_recurrent_2010}, contextualized speech representations can be learned by predicting future values of a simple representation \cite{oord_representation_2018, chung_unsupervised_2019, schneider_wav2vec_2019, chung_generative_2020, jiang_further_2021} (\textbf{\textsc{prd}}, table \ref{tab:model-taxonomy}). Modeling spectrograms directly, autoregressive predictive coding (APC, \citealp{chung_unsupervised_2019}) is perhaps the simplest example in this category. The forward pass and loss are computed as
\begin{align}
    \mathbf{c}_{t} &= f(\mathbf{x}_{1:t})  \label{eq_brief:apc_f} \\ 
    \hat{\mathbf{x}}_{t+k} &= g(\mathbf{c}_{t}) \\
    \mathcal{L}_t &= \lVert \hat{\mathbf{x}}_{t+k} - \mathbf{x}_{t+k} \rVert_1\enspace.
\end{align}
%
\noindent Here, $f(\cdot)$ and $g(\cdot)$ are parameterized by neural networks such that each $\mathbf{c}_t$ is only conditioned on previous inputs $\mathbf{x}_{1:t}$ and $\hat{\mathbf{x}}_{t+k}$ is computed step-wise. \citet{chung_unsupervised_2019} use a stack of unidirectional LSTMs for $f(\cdot)$ and a linear regression layer for $g(\cdot)$. Tasks that seek to predict or reconstruct the input are very common. In the literature, these are often jointly referred to as reconstruction tasks (\textbf{\textsc{rec}}, table \ref{tab:model-taxonomy}) \cite{liu_tera_2021, wang_unispeech_2021}, although this is somewhat misleading in the case of prediction.

Contrary to generative models, such as WaveNet \cite{oord_wavenet_2016}, the APC model is not restricted to next-step prediction. Instead, it predicts $k > 0$ steps ahead in order to ensure that the model does not learn a trivial solution by exploiting the smoothness of the signal. Depending on the downstream task, we are often interested in learning so-called slow features that will typically span multiple input frames \cite{wiskott_slow_2002}. Even the smallest linguistic units of speech, phonemes, tend to span $0.1$ seconds on average \cite{garofolo_timit_1993}, whereas spectrogram frames $\mathbf{x}_t$ are typically computed at $0.01$ second intervals. However, sometimes local smoothness is explicitly used to define the task \cite{badino_autoencoder_2014, jati_speaker2vec_2017, jati_neural_2019}.

\paragraph{Contrastive models}
Speech contains localized noise (e.g., phase shifts) that does not inform slow feature learning. Thus, directly modeling speech might not be the best way to learn contextualized representations. Contrastive predictive coding (CPC,  \citealp{oord_representation_2018}) targets a local variable $\mathbf{v}_{1:T}$, learned from the model input $\mathbf{x}_{1:T}$, instead of the input itself. The forward pass is
%
\begin{align}
    \mathbf{v}_{t} &= f_{v}(\mathbf{x}_{t-r:t+r}) \label{eq_brief: cpc local representation} \\
    \mathbf{c}_{t} &= f_{c}(\mathbf{v}_{1:t}) \\
    \hat{\mathbf{v}}_{t,k} &= g_k(\mathbf{c}_{t}) \enspace,
\end{align}
%
\noindent where $f_v(\cdot)$ is a convolutional neural network, such that each $\mathbf{v}_{t}$ only encodes information from a limited receptive field $2r+1$. Again, $f_c(\cdot)$ should be limited to condition each $\mathbf{c}_t$ on previous time-steps $\mathbf{v}_{1:t}$ and $g_k(\cdot)$ is a step-wise transformation. The loss is based on noise constrastive estimation \cite{gutmann_noisecontrastive_2010} and is given by
\begin{align}
    \mathcal{L}_{t,k} &= - \log \left(\frac{\exp(\hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{t+k})}{\sum_{n \sim \mathcal{D}} \exp(\hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{n})} \right)\enspace .
    \label{eq_brief: cpc loss}
\end{align}
%
\noindent Here, $\mathcal{D}$ is a set of indices including the target index $t+k$ and negative samples drawn from a proposal distribution, which is typically taken to be a uniform distribution over the set $\{1,\dots,T\}$. Note that the loss is also indexed by $k$ to show that CPC targets multiple offsets. The APC model is easily extended in a similar way \cite{chung_improved_2020}.

@@

A model closely related to the original CPC model is wav2vec \cite{schneider_wav2vec_2019}. It uses a different parameterization of the functions $f_v(\cdot)$ and $f_c(\cdot)$, and modifies the loss to consider a binary prediction task, such that we have
%
\begin{align}
    \mathcal{L}_{t,k} &= - \log(\sigma(\hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{t+k})) - \sum_{n \sim \mathcal{D}} \log(\sigma(-\hat{\mathbf{v}}_{t,k}^{\text{\tiny T}}\mathbf{v}_{n})) \,.
    \label{eq_brief: wav2vec loss}
\end{align}

@@

\paragraph{Masking-based models} One downside of predictive tasks is that models are primarily unidirectional. Some work has extended APC and CPC inspired models with separate encoders operating in opposite directions \cite{ling_deep_2020, kawakami_learning_2020, borgholt_scaling_2021}, but these models are still restricted to process left and right context separately. Inspired by the masked language model task used for text-based representation learning \cite{devlin_bert_2018}, several papers have used masking to overcome this challenge (\textbf{\textsc{msk}}, table \ref{tab:model-taxonomy}). Masking refers to replacing parts of the input with zeros or a learned masking vector. For zero-masking \cite{jiang_improving_2019, liu_mockingjay_2020, wang_unsupervised_2020, chi_audio_2020, ling_decoar_2020}, we have
\begin{align}
    \mathbf{c}_{1:T} &= f(\mathbf{x}_{1:T} \circ \mathbf{m}_{1:T})\\
    \mathbf{\hat{x}}_{t} &= g(\mathbf{c}_{t}) \\
    \mathcal{L}_t &= \lVert \mathbf{\hat{x}}_{t} - \mathbf{x}_{t} \rVert_1\enspace,
\end{align}
where the $\circ$ operator denotes the Hadamard product, $f(\cdot)$ is typically a transformer encoder or a bidirectional recurrent neural network, $g(\cdot)$ is a step-wise transformation, and $\mathbf{m}_{1:T}$ is a mask such that $m_{t,i} \in \{0,1\}$. Alternatively, $\mathbf{m}_{1:T}$ is used to select which $\mathbf{x}_t$ are replaced by a learned masking vector. The entries of $\mathbf{m}_{1:T}$ are determined by some stochastic policy. One frequent inspiration is SpecAugment \cite{park_specaugment_2019}, which was originally proposed for supervised speech recognition and applies frequency and time masking to spectrogram representations.
While temporal masking is most common, frequency masking has also been adopted for representation learning  \cite{wang_unsupervised_2020}. A simple, yet popular, masking strategy is to draw a proportion of input indices $t_i\sim\{1,\dots,T-M\}$ without replacement, and then mask $\{t_i, \dots, t_i+M\}$ \cite{baevski_wav2vec_2020, hsu_hubert_2021, ling_decoar_2020}.

Combining masking with a contrastive loss, wav2vec 2.0 was the first work to show that a competitive speech recognition model can be learned by fine-tuning a pre-trained model with as little as 10 minutes of labeled data. For this model
%
\begin{align}
    \mathbf{v}_{t} &= f_v(\mathbf{x}_{t-r:t+r}) \\
    \mathbf{c}_{1:T} &= f_c(\mathbf{v}_{1:T} \circ \mathbf{m}_{1:T}) \\
    \mathbf{q}_t &= g_q(\mathbf{v}_t) \enspace . \label{eq_brief: w2v2 qtz}
\end{align}

\noindent Here, $f_v(\cdot)$ is a convolutional neural network, $f_c(\cdot)$ is a transformer encoder \cite{vaswani_attention_2017} and $g_q(\cdot)$ is a quantization module used to learn targets from the localized variable $\mathbf{v}_{1:T}$. Computing quantized targets this way requires an extra loss term, which we will present when we discuss quantization in general below. The contrastive loss for wav2vec 2.0 is similar to that of the CPC model,
%
\begin{align}
    \mathcal{L}_t &= - \log \left(\frac{\exp(S_{\text{c}}(\mathbf{c}_{t}, \mathbf{q}_{t}))}{\sum_{n \sim \mathcal{D}} \exp(S_{\text{c}}(\mathbf{c}_{t}, \mathbf{q}_{n}))} \right) \enspace , \label{eq_brief: w2v2 loss}
\end{align}
%
\noindent where $S_{\text{c}}(\cdot)$ is the cosine similarity and the negative samples in $\mathcal{D}$ are sampled from other masked time-steps.

@@

\noindent Non-autoregressive predictive coding (NPC, \citealp{liu_nonautoregressive_2020}) tries to resolve this by using a convolutional neural network where the kernel is masked instead of the input.
This allows for complete data utilization, but limits the amount of context encoded in the learned representation.
Figure \ref{fig:ssl_grid} summarizes the models discussed so far.

\begin{figure}[!t]
    \centering
    \setlength\tabcolsep{1.5pt}
    \begin{tabular}{>{\centering\arraybackslash} m{4mm}  >{\centering\arraybackslash} m{0.40\textwidth}|>{\centering\arraybackslash} m{0.40\textwidth}}
          & {\small \textsc{predictive}} & {\small \textsc{masking-based}} \\
        \rotatebox{90}{{\small \textsc{reconstruct}}} & \includegraphics[width=0.40\textwidth]{paper_brief/REC_PRD.pdf} & \includegraphics[width=0.40\textwidth]{paper_brief/REC_MSK.pdf}  \\
        \midrule
        \rotatebox{90}{{\small \textsc{contrastive}}} & \includegraphics[width=0.40\textwidth]{paper_brief/CON_PRD.pdf} & \includegraphics[width=0.40\textwidth]{paper_brief/CON_MSK.pdf}
    \end{tabular}
    \caption{Schematic of self-supervised methods. Each subfigure illustrates the loss computation for a single time-step. The temporal subscript has been left out for simplicity.}
    \label{fig:ssl_grid}
\end{figure}

@@

            \midrule
            \multicolumn{12}{c}{\textsc{Self-supervised models}} \\
            \midrule
            %                                                           MSK      PRD      CON      REC      QTZ      GEN      LOC      GLB     VAR      FRZ      FTN
            \textbf{Audio Word2vec} \footnotesize\cite{chung_audio_2016}      & 2016 Mar. & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark \\
            \textbf{Speech2Vec} \footnotesize\cite{chung_speech2vec_2018}     & 2018 Mar. & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark \\
            \textbf{Unspeech} \footnotesize\cite{milde_unspeech_2018}         & 2018 Apr. & \xmark & \cmark & \cmark & \xmark & \xmark & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark \\

@@

The inference model may also be used to infer a global latent variable (\textbf{\textsc{glb}}, table \ref{tab:lvm-model-primitives}) that might encode global information about $\mathbf{x}$. 
While it must be included in the prior model it might not be in the observation model, if the model also has a local latent variable. 
Finally, latent variables are often made to depend autoregressively on past inferred values, e.g. $q(\mathbf{z}_t|\mathbf{z}_{1:t-1},\mathbf{x}_{1:t})$ (\textbf{\textsc{arz}}, table \ref{tab:lvm-model-primitives}) \cite{chung_recurrent_2015, fraccaro_sequential_2016}.

\paragraph{Multiscale and hierarchical models} 
Some work has explored using a hierarchy of latent variables (\textbf{\textsc{hie}}, table \ref{tab:lvm-taxonomy}). This allows encoding the inductive bias that speech contains information at different temporal scales by letting the latent variables operate at different temporal scales \cite{hsu_unsupervised_2017}. \citet{khurana_factorial_2019} propose using a temporal latent variable along with a global latent variable. 
Recent work has focused on learning a deeper latent hierarchy with five latent variables \cite{aksan_stcn_2019}.
